### 2015-07-21 tex/th4.tex

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
\item \input{case1.tex}

\item \input{case2part1.tex}
\input{lemmafirst.tex}
\input{case2general.tex}
\input{case2generalproof.tex}

\item \input{case3.tex}
\end{enumerate}
\end{proof}