### 2016-04-06 tex/below.tex
'5abr'

\section{Counting classes below \#P}

We define $\totp$ using the following function. For each non-deterministic Turing machine $M$,
\[
	\tot_M(x) = (\text{\#paths of $M$ on input $x$}) - 1.
\]
A function $f:\{0,1\}^*\to\N$ is in $\totp$ if there exists a polynomial-time non-deterministic Turing machine $M$ such that $f = \tot_M$.

We define $\spanl$ using the following function. For each non-deterministic Turing machine $M$,
\[
	\sp_M(x) = \vert\{w \mid \text{$w$ is an output of $M$ on input $w$}\}\vert.
\]
A function $f:\{0,1\}^*\to\N$ is in $\spanl$ if there exists a log-space non-deterministic Turing machine $M$ such that $f = \sp_M$.

\begin{theo}
	$\spanl \subseteq \totp$.
\end{theo}
\begin{proof}
	Let $f:\{0,1\}^*\to\N$ be a function in $\spanl$ and let $M$ be a non-deterministic log-space Turing machine such that $f = \sp_M$. We will provide a polynomial-time non-deterministic Turing machine $M'$ such that for each $x\in\{0,1\}^*$ it holds that $\sp_M(x) = \tot_{M'}(x)$. First we note that given a configuration $s$ of $M$ on input $x$ checking whether there exists a final state of $M$ which is reachable from $s$ can be done in polynomial time. This implies that checking whether $f(x) > 0$ can also be done in polynomial time. We present this machine as a non-deterministic algorithm.

	\begin{algorithm}
		\caption{Calculate $\sp_M(x)$}
		\label{alg1}
		\begin{algorithmic}
			\REQUIRE $n \geq 0 \vee x \neq 0$
			\ENSURE $y = x^n$
			\STATE $y \leftarrow 1$
			\IF{$n < 0$}
			\STATE $X \leftarrow 1 / x$
			\STATE $N \leftarrow -n$
			\ELSE
			\STATE $X \leftarrow x$
			\STATE $N \leftarrow n$
			\ENDIF
			\WHILE{$N \neq 0$}
			\IF{$N$ is even}
			\STATE $X \leftarrow X \times X$
			\STATE $N \leftarrow N / 2$
			\ELSE[$N$ is odd]
			\STATE $y \leftarrow y \times X$
			\STATE $N \leftarrow N - 1$
			\ENDIF
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}

\end{proof}