### 2016-12-27 LICS/syntactic.tex

%!TEX root = main.tex

In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$. We show that by restricting $\LL$ we can find different subclasses below $\shp$ with interesting computational and closure properties.

\subsection{Counting hierarchy below $\shp$}
\input{saluja}

\subsection{Horn Counting Classes}
\input{horn-counting}