### 2017-03-03 LMCS/syntactic.tex

%!TEX root = main.tex

\subsection{The $\eqso(\fo)$-hierarchy inside $\shp$}
\label{sec-hier-shp}
\input{saluja}

\subsection{Defining a class of functions with easy decision versions and good closure properties}
\label{sec-clo}
\input{syntactic-clo}

\subsection{Defining a class of functions with easy decision versions and natural complete problems}
\label{sec-horn}
\input{syntactic-horn}