### 2017-03-23 LMCS/proof-capture-fpspace-poly.tex

For the first condition, [analogous to 4.4]

For the second condition, [analogous to 4.4]