### 2017-03-27 LMCS/proof-capture-fpspace-poly.tex

In this proof, both conditions are analogous to theorem \ref{theo:capture-fp}. For condition (1), each $\pfp$ formula $\varphi$ can be evaluated in $\pspace$. The mentioned closure properties are also possessed by $\nfpspace$. For condition (2) we define the same variables and the same language $L$, which in this case is in $\pspace$. The same construction of $\alpha$, which in turn is in $\qfo(\pfp)$, can be used to show that the condition holds.